---
title: "Data Analysis"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
library(tidyverse)
library(knitr)
library(dplyr)
library(lubridate)
knitr::opts_chunk$set(echo = TRUE)
```

##### Has there been an increase in the rate of requests made (Per week) ever since the implementation of the new feature? 


##### Is there a difference between the rate of increase between `professors` and `employers`? What does this mean?

### Data Analysis

```{r}
requests_dat <- read_csv("data/requests_Mar18_2019.csv")
educators <- read_csv("data/educatoraccounts.csv")
employers <- read_csv("data/employeraccounts.csv")

employers <- employers %>% select(Id, Type)
colnames(employers) <- c("Actor Id", "Type")

## `requestable model == project` means an `employer` is requesting
## `requestable model == course` means an `employer` is requesting
## There are some valid requests on August 30, 2018, so to make sure they are not deleted we remove all observations where`actor id` is NA.
## Filtered out observations with data "2018-04-12" as there were over 250 observations on that that. It is highly unlikely that 250 requests were made it one day. We assume its some sort of data migration
requests_filtered <- requests_dat %>% mutate(type=ifelse(`Requestable Model`=="course", "professor","employer")) %>% rename(date_created=`Day of Created At`, date_updated=`Day of Updated At`, date_expired=`Day of Expired At`) %>% mutate(date_created=dmy(date_created)) %>% mutate(date_expired=dmy(date_expired)) %>% mutate(date_updated=dmy(date_updated)) %>% filter(!is.na(`Actor Id`)) %>%
  filter(date_created>min(date_created))

main_requests <- requests_filtered

## data frame that stores the number of requests by week
requests <- requests_filtered %>% group_by(date_created = floor_date(date_created, unit="day")) %>% summarize(num_requests=n()) %>% complete(date_created = seq.Date(min(date_created), max(date_created), by="day")) %>% mutate(num_requests=(ifelse(is.na(num_requests), 0, num_requests)))


## filtered_requests removes the outlier on "2019-02-28" that had 50 requests.
filtered_requests <- requests %>% filter(date_created != "2019-02-28") %>% mutate(state=ifelse(date_created<"2018-11-21","before", "after"))
```
## data frame that stores the number of requests by week
requests <- requests %>% group_by(date_created = floor_date(date_created, unit="day")) %>% summarize(num_requests=n()) %>% complete(date_created = seq.Date(min(date_created), max(date_created), by="day")) %>% mutate(num_requests=(ifelse(is.na(num_requests), 0, num_requests)))

1. State hypotheses:
Null hypothesis (H0): the request acceptance rate is not influenced by the addition of the “request expiry” feature added on November 21st, 2018.

Alternative hypothesis (HA): the request acceptance rate is influenced by the addition of the “request expiry” feature added on November 21st, 2018.

2. Calculate Test Statistic:
```{r}
accepted_before <- length(filter(main_requests, date_created < "2018-11-21" & State == "accepted")$State)
non_accepted_before <- length(filter(main_requests, date_created < "2018-11-21" & State != "accepted")$State)

accepted_after <- length(filter(main_requests, date_created >= "2018-11-21" & State == "accepted")$State)
non_accepted_after <- length(filter(main_requests, date_created >= "2018-11-21" & State != "accepted")$State)

time <- c("before new feature", "after new feature")
num_accept <- c(accepted_before, accepted_after)
num_non_accept <- c(non_accepted_before, non_accepted_after)

feature_data <- data.frame(time, "accepted" = num_accept, "non-accepted" = num_non_accept)

feature_data

#calculate counts
n_before <- main_requests %>% filter(date_created < "2018-11-21") %>% summarise(n())
n_after <- main_requests %>% filter(date_created >= "2018-11-21") %>% summarise(n())

p_diff <- accepted_after/n_after - accepted_before/n_before
as.numeric(p_diff)
acceptance_rate = 272/272
acceptance_rate
nonacceptance_rate = 237/249
nonacceptance_rate
```
3. Simulate Test Statistic under H0
```{r}
set.seed(130)
repetitions <- 500
simulated_stats <- rep(NA, repetitions)
n_before <- main_requests %>% filter(date_created < "2018-11-21") %>% summarise(n())
n_after <- main_requests %>% filter(date_created >= "2018-11-21") %>% summarise(n())

accepted_after <- main_requests %>% filter(date_created >= "2018-11-21" & State=="accepted") %>% summarize(n()) 

accepted_before <- main_requests %>% filter(date_created < "2018-11-21" & State=="accepted") %>% summarize(n())

test_stat <- as.numeric(accepted_after/n_after - accepted_before/n_before)

for (i in 1:repetitions)
{
  sim <- main_requests %>% mutate(State = sample(State))
  accepted_after <- sim %>% filter(State=="accepted" & date_created >= "2018-11-21") %>% summarize(n())
  accepted_before <- sim %>% filter(State=="accepted" & date_created < "2018-11-21") %>% summarize(n())
  p_diff <- accepted_after/n_after - accepted_before/n_before
  simulated_stats[i] <- as.numeric(p_diff)
}

sim <- data_frame(p_diff=simulated_stats)

ggplot(sim, aes(x=p_diff)) + geom_histogram(binwidth=0.01)

ggplot(sim, aes(p_diff)) + geom_histogram(binwidth=0.01) + geom_vline(xintercept = abs(test_stat), color="red") + geom_vline(xintercept = -1*abs(test_stat), color="red") + labs(x = "Simulated difference in proportion accepted before and after new feature is added") 

```

4. Assess Evidence against H0: P-value
```{r}
extreme_count <- sim %>% filter(p_diff >= abs(test_stat) | p_diff <= -1*abs(test_stat)) %>% summarise(n())
as.numeric(extreme_count)

p_value <- as.numeric(extreme_count)/repetitions
as.numeric(p_value)
```

5. Make conclusion
Since we calculated an extremely large p-value of 0.726, we can conclude that there is a almost no amount of evidence against the null hypothesis. Furthermore, because a rate of 1:1 acceptance to non-acceptance rate only changed to a non-acceptance rate of 0.95:1 after the feature was added, we can conclude that the acceptance rate was not changed much by the feature being added. 

6. Furthermore
Because of our ambigouous results, we further investigate the State variable in requests to compare expiry rates before and after the feature is added. However, because requests do not exist with the state "expired" before the feature is added on November 21, 2018, we consider any requests that have a difference of date_updated and date_created larger than 14 days as "would be" expired data.

```{r}
#new_requests <- data.frame(Date=as.Date(character()),
#                 File=character(), 
#                 User=character(), 
#                stringsAsFactors=FALSE) 

#for (row in 1:nrow(requests)) {
  
#  month_created <- as.numeric(substr(date_created, 6, 7))
#  month_updated <- as.numeric(substr(date_updated, 6, 7))

#  day_created <- as.numeric(substr(date_created, 9, 10))
#  day_updated <- as.numeric(substr(date_updated, 9, 10))
  
#  if(month_created == month_updated & (day_updated - day_created) > 14) {
#  rownames(rbind(t(requests[row,]),t(new_requests))
#  }
#}

n_before <- main_requests %>% filter(date_created < "2018-11-21" & date_expired != NA)%>% summarise(n())
n_after <- main_requests %>% filter(date_created >= "2018-11-21") %>% summarise(n())
expired_before <- as.numeric(filter(main_requests, date_created < "2018-11-21") %>% summarise(n()))
expired_after <- as.numeric(filter(main_requests, date_created >= "2018-11-21" & State == "expired") %>% summarise(n()))

before_ratio <- expired_before / n_before * 100
after_ratio <- expired_after / n_after * 100
  
expiry_df <- data.frame(
  group = c("Before", "After"),
  value = c(expired_before, expired_after)
  )
expiry_df

ggplot(data=expiry_df, aes(x=group, y = value)) +
  geom_bar(stat="identity", width=0.5)
```

